Summary of the things that I learned

### 2024-11-14 200-category-theory.tex
Improve titles, make some sentences active instead of passive

\subsection{Adjoint equivalences}
An (adjoint) equivalence of categories has two equivalent definitions, that only differ in which notion they take as the base. In this case, we will use the definition that is based on adjunctions:
\begin{definition}\label{def:equivalence-of-categories}
  An \iindex{adjoint equivalence} between categories $ \C $ and $ \D $ is a pair of adjoint functors $ L \dashv R $ like above such that the unit $ \eta: \id{\C} \Rightarrow L \bullet R $ and co-unit $ \epsilon: R \bullet L \Rightarrow \id{\D} $ are isomorphisms of functors.
\end{definition}

@@

\subsection{Weak equivalences}
There is also the notion of `weak equivalence'. In some cases, this is equivalent to an adjoint equivalence (for example, when its domain is univalent, see Section \ref{sec:univalence-principle}).
\begin{definition}
  A functor $ F: \C \to \D $ is called a \iindex{weak equivalence} if it is essentially surjective and fully faithful.
\end{definition}

\subsection{Exponential objects}
Note that in the category of sets, for all $ X, Y: \SET $, we have a set of functions $ (X \to Y) $. Also, for all $ X, Y, Z $, there is a (natural) bijection
\[ (X \times Y \to Z) \cong (X \to (Y \to Z)) \]
which we can also write as
\[ \SET(X \times Y, Z) \cong \SET(X, (Y \to Z)). \]
In other words, we have functors $ X \mapsto X \times Y $ and $ Z \mapsto (Y \to Z) $, and these two form an adjunction. The following generalizes this
\begin{definition}
  A category $ \C $ has \textit{exponential objects}\index{exponential objects} (or \textit{exponentials}) if for all $ X: \C $, the functor $ X^\prime \mapsto X^\prime \times X $ has a right adjoint, which we denote $ Y \mapsto Y^X $.
\end{definition}

@@

\subsection{Forgetful functors and free objects}

@@

\subsection{Extension and restriction of scalars}

>>>>

\subsection{Adjoint Equivalences}
An (adjoint) equivalence of categories has two equivalent definitions, that only differ in which notion they take as the base. In this case, we will use the definition that is based on adjunctions:
\begin{definition}\label{def:equivalence-of-categories}
  An \iindex{adjoint equivalence} between categories $ \C $ and $ \D $ is a pair of adjoint functors $ L \dashv R $ like above such that the unit $ \eta: \id{\C} \Rightarrow L \bullet R $ and co-unit $ \epsilon: R \bullet L \Rightarrow \id{\D} $ are isomorphisms of functors.
\end{definition}

@@

\subsection{Weak Equivalences}
There is also the notion of `weak equivalence'. In some cases, this is equivalent to an adjoint equivalence (for example, when its domain is univalent, see Section \ref{sec:univalence-principle}).
\begin{definition}
  A functor $ F: \C \to \D $ is called a \iindex{weak equivalence} if it is essentially surjective and fully faithful.
\end{definition}

\subsection{Exponential Objects}
Note that in the category of sets, for all $ X, Y: \SET $, we have a set of functions $ (X \to Y) $. Also, for all $ X, Y, Z $, there is a (natural) bijection
\[ (X \times Y \to Z) \cong (X \to (Y \to Z)) \]
which we can also write as
\[ \SET(X \times Y, Z) \cong \SET(X, (Y \to Z)). \]
In other words, we have functors $ X \mapsto X \times Y $ and $ Z \mapsto (Y \to Z) $, and these two form an adjunction. The following generalizes this
\begin{definition}
  A category $ \C $ has \textit{exponential objects}\index{exponential objects} (or \textit{exponentials}) if for all $ X: \C $, the functor $ X^\prime \mapsto X^\prime \times X $ has a right adjoint, which we denote $ Y \mapsto Y^X $.
\end{definition}

@@

\subsection{Forgetful Functors and Free Objects}

@@

\subsection{Extension and Restriction of Scalars}

### 2024-11-14 400-algebraic-structures.tex
Improve titles, make some sentences active instead of passive

\section{\texorpdfstring{$ \lambda $-}{lambda-}theories}\label{sec:lambda-theories}

@@

\subsection{The $ \lambda $-calculus operations}\label{subsec:lambda-calculus-operations}
For a $ \lambda $-theory $ L $, we have variables $ x_{n, i} : L_n $ and $ \lambda $-abstraction $ f \mapsto \lambda(f) $. We will sometimes denote $ \lambda(f) $ as $ \lambda x_{n + 1}, f $ for $ f: L_{n + 1} $. Now, consider the element $ \rho(x_{1, 1}) : L_2 $.
\begin{definition}[\coqident{AlgebraicTheories.LambdaTheories}{app'}]
  Using the substitution, we have binary operations on the $ L_n $, sending $ (f, g) : L_n \times L_n $ to $ \rho(x_{1, 1}) \bullet (f, g) : L_n $. We will denote $ \rho(x_{1, 1}) \bullet (f, g) $ as $ f g $, and this gives us our application operation.
\end{definition}
This means that we can interpret all three operations (\texttt{var}, \texttt{app} and \texttt{abs}) of the $ \lambda $-calculus in $ L $.

@@

\subsection{The free algebraic theory on a set}
\begin{example}[\coqident{AlgebraicTheories.Examples.FreeTheory}{free_functor}]
  Let $ S $ be a set. We can construct an algebraic theory $ F(S) $ by taking $ F(S)_n = S \sqcup \{ 1, \dots, n \} $ with projections $ x_i = i $ and substitution
  \begin{align*}
    i \bullet g &= g_i & s \bullet g &= s
  \end{align*}
  for $ i : \{ 1, \dots, n \} $ and $ s : S $.

@@

\subsection{The free \texorpdfstring{$ \lambda $}{lambda}-theory on a set}

@@

\subsubsection{About $ \Lambda $-algebra morphisms}

@@

\subsection{The free object algebraic theory}

@@

\subsection{The terminal theory}
\begin{example}[\coqident{AlgebraicTheories.Examples.OnePointTheory}{one_point_theory}]
  We can create the trivial algebraic theory $ T $ by taking $ T_n = \{ \star \} $, with projections $ x_i = \star $ and substitution $ \star \bullet \star = \star $. Taking $ \lambda(\star) = \star $ and $ \rho(\star) = \star $, we give it a $ \lambda $-theory structure (with $ \beta $ and $ \eta $-equality). Checking that this is indeed an algebraic theory and even a $ \lambda $-theory is trivial.

@@

\subsection{The endomorphism theory}

@@

\subsection{The theory algebra}
\begin{example}[\coqident{AlgebraicTheories.Examples.TheoryAlgebra}{theory_algebra}]\label{ex:theory-algebra}
  Let $ T $ be an algebraic theory and $ n $ a natural number. We can endow the $ T_n $ with a $ T $-algebra structure, by taking the substitution operator of $ T $ as the $ T $-action. Since this commutes with the substitution operator and the projections, $ T_n $ is a $ T $-algebra.
\end{example}

@@

\subsection{The initial presheaf}
\begin{example}
  Let $ T $ be an algebraic theory. We can construct a $ T $-presheaf $ P $, with $ P_n = \emptyset $. Then $ \bullet : P_m \times T_n^m \to P_m $ is trivial, and the presheaf laws hold trivially.
\end{example}

@@

\subsection{The theory presheaf}
\begin{example}[\coqident{AlgebraicTheories.Presheaves}{theory_presheaf}]
  Let $ T $ be an algebraic theory. We can endow $ T $ with a $ T $-presheaf structure, by taking the substitution operator of $ T $ as the action on $ T $. Since this commutes with the substitution operator and the projections, $ T $ is a $ T $-presheaf.
\end{example}

@@

\subsection{The ``+\texorpdfstring{$ \ell $}{l}'' presheaf}

\begin{example}[\coqident{AlgebraicTheories.Examples.Plus1Presheaf}{plus_1_presheaf}]
  Given a $ T $-presheaf $ Q $, we can construct a presheaf $ A(Q, \ell) $ with $ A(Q, \ell)_n = Q_{n + \ell} $ and, for $ q: A(Q, \ell)_m $ and $ f: T_n^m $, action
  \[ q \bullet_{A(Q, \ell)} f = q \bullet_Q ((\iota_{n, \ell} (f_i))_i + (x_{n + i})_i). \]
\end{example}

\begin{lemma}[\coqident{AlgebraicTheories.RepresentationTheorem}{theory_presheaf_exponentiable}]
  For all $ m $ and all $ T $-presheaves $ Q $, $ A(Q, m) $ is the exponential object $ Q^{T^m} $.
\end{lemma}
\begin{proof}
  We will show that $ A(-, m) $ constitutes a right adjoint to the functor $ - \times T^m $. We will do this using universal arrows.

  For $ Q $ a $ T $-presheaf, take the arrow $ \varphi: A(Q, m) \times T^m \to Q $ given by $ \varphi(q, t) = q \bullet_Q ((x_{n, i})_i + t) $ for $ q: A(Q, m)_n = Q_{n + m} $ and $ t: T^m_n $.

  Now, given a $ T $-presheaf $ Q^\prime $ and a morphism $ \psi: Q^\prime \times T^m \to Q $. Define $ \tilde \psi: Q^\prime_n \to A(Q, m)_n $ by $ \tilde \psi(q) = \psi(\iota_{n, m}(q), (x_{n + i})_i) $.

  Then $ \psi $ factors as $ (\tilde \psi \times \id{T^m}) \cdot \varphi $. Also, some equational reasoning shows that $ \tilde \psi $ is unique, which proves that $ \varphi $ indeed is a universal arrow.
\end{proof}

>>>>

\section{\texorpdfstring{$ \lambda $-}{lambda-}Theories}\label{sec:lambda-theories}

@@

\subsection{The $ \lambda $-Calculus Operations}\label{subsec:lambda-calculus-operations}
For a $ \lambda $-theory $ L $, we have variables $ x_{n, i} : L_n $ and $ \lambda $-abstraction $ f \mapsto \lambda(f) $. We will sometimes denote $ \lambda(f) $ as $ \lambda x_{n + 1}, f $ for $ f: L_{n + 1} $. Now, consider the element $ \rho(x_{1, 1}) : L_2 $.
\begin{definition}[\coqident{AlgebraicTheories.LambdaTheories}{app'}]
  Using the substitution, we have binary operations on the $ L_n $, sending $ (f, g) : L_n \times L_n $ to $ \rho(x_{1, 1}) \bullet (f, g) : L_n $. We will denote $ \rho(x_{1, 1}) \bullet (f, g) $ as $ f g $, and this gives us our application operation.
\end{definition}
This means that we can interpret all three operations (\texttt{var}, \texttt{app} and \texttt{abs}) of the $ \lambda $-calculus in $ L $.

@@

\subsection{The Free Algebraic Theory on a Set}
\begin{example}[\coqident{AlgebraicTheories.Examples.FreeTheory}{free_functor}]
  Let $ S $ be a set. We can construct an algebraic theory $ F(S) $ by taking $ F(S)_n = S \sqcup \{ 1, \dots, n \} $ with projections $ x_i = i $ and substitution
  \begin{align*}
    i \bullet g &= g_i & s \bullet g &= s
  \end{align*}
  for $ i : \{ 1, \dots, n \} $ and $ s : S $.

@@

\subsection{The Free \texorpdfstring{$ \lambda $}{lambda}-Theory on a Set}

@@

\subsubsection{About $ \Lambda $-Algebra Morphisms}

@@

\subsection{The Free Object Algebraic Theory}

@@

\subsection{The Terminal Theory}
\begin{example}[\coqident{AlgebraicTheories.Examples.OnePointTheory}{one_point_theory}]
  We can create the trivial algebraic theory $ T $ by taking $ T_n = \{ \star \} $, with projections $ x_i = \star $ and substitution $ \star \bullet \star = \star $. Taking $ \lambda(\star) = \star $ and $ \rho(\star) = \star $, we give it a $ \lambda $-theory structure (with $ \beta $ and $ \eta $-equality). Checking that this is indeed an algebraic theory and even a $ \lambda $-theory is trivial.

@@

\subsection{The Endomorphism Theory}

@@

\subsection{The Theory Algebra}
\begin{example}[\coqident{AlgebraicTheories.Examples.TheoryAlgebra}{theory_algebra}]\label{ex:theory-algebra}
  Let $ T $ be an algebraic theory and $ n $ a natural number. We can endow the $ T_n $ with a $ T $-algebra structure, by taking the substitution operator of $ T $ as the $ T $-action. Since this commutes with the substitution operator and the projections, $ T_n $ is a $ T $-algebra.
\end{example}

@@

\subsection{The Initial Presheaf}
\begin{example}
  Let $ T $ be an algebraic theory. We can construct a $ T $-presheaf $ P $, with $ P_n = \emptyset $. Then $ \bullet : P_m \times T_n^m \to P_m $ is trivial, and the presheaf laws hold trivially.
\end{example}

@@

\subsection{The Theory Presheaf}
\begin{example}[\coqident{AlgebraicTheories.Presheaves}{theory_presheaf}]
  Let $ T $ be an algebraic theory. We can endow $ T $ with a $ T $-presheaf structure, by taking the substitution operator of $ T $ as the action on $ T $. Since this commutes with the substitution operator and the projections, $ T $ is a $ T $-presheaf.
\end{example}

@@

\subsection{The `\texorpdfstring{$ n + p $}{n + p}'-Presheaf}

\begin{example}[\coqident{AlgebraicTheories.Examples.Plus1Presheaf}{plus_1_presheaf}]
  Given a $ T $-presheaf $ Q $, we can construct a presheaf $ A(Q, p) $ with $ A(Q, p)_n = Q_{n + p} $ and, for $ q: A(Q, p)_m $ and $ f: T_n^m $, action
  \[ q \bullet_{A(Q, p)} f = q \bullet_Q ((\iota_{n, p} (f_i))_i + (x_{n + i})_i). \]
\end{example}

\begin{lemma}[\coqident{AlgebraicTheories.RepresentationTheorem}{theory_presheaf_exponentiable}]
  For all $ p $ and all $ T $-presheaves $ Q $, $ A(Q, p) $ is the exponential object $ Q^{T^p} $.
\end{lemma}
\begin{proof}
  We will show that $ A(-, p) $ constitutes a right adjoint to the functor $ - \times T^p $. We will do this using universal arrows.

  For $ Q $ a $ T $-presheaf, take the arrow $ \varphi: A(Q, p) \times T^p \to Q $ given by $ \varphi(q, t) = q \bullet_Q ((x_{n, i})_i + t) $ for $ q: A(Q, p)_n = Q_{n + p} $ and $ t: T^p_n $.

  Now, given a $ T $-presheaf $ Q^\prime $ and a morphism $ \psi: Q^\prime \times T^p \to Q $. Define $ \tilde \psi: Q^\prime_n \to A(Q, p)_n $ by $ \tilde \psi(q) = \psi(\iota_{n, p}(q), (x_{n + i})_i) $.

  Then $ \psi $ factors as $ (\tilde \psi \times \id{T^p}) \cdot \varphi $. Also, some equational reasoning shows that $ \tilde \psi $ is unique, which proves that $ \varphi $ indeed is a universal arrow.
\end{proof}

### 2024-11-14 500-previous-work.tex
Improve titles, make some sentences active instead of passive

\subsection{Taylor's proof}

@@

\section{The \texorpdfstring{$ \lambda $}{lambda}-calculus is Algebraic}\label{sec:selinger}
As we mentioned in Remark \ref{rem:lambda-theory-or-algebra}, using tools from universal algebra, there are two approaches to study $ \lambda $-calculus-like structures. One can either study $ \lambda $-theories or algebras for the algebraic theory $ \Lambda $. We will see that Hyland shows that these are equivalent.

>>>>

\subsection{Taylor's Proof}

@@

\section{The \texorpdfstring{$ \lambda $}{lambda}-Calculus is Algebraic}\label{sec:selinger}
As we mentioned in Remark \ref{rem:lambda-theory-or-algebra}, using tools from universal algebra, there are two approaches to study $ \lambda $-calculus-like structures. One can either study $ \lambda $-theories or algebras for the algebraic theory $ \Lambda $. We will see that Hyland shows that these are equivalent.
